Remote sensing riverine gas exchange from SWOT and implications for estimating carbon efflux from rivers

### 2021-02-12 manuscript_v1.Rmd
message

---
title: "Remote sensing the gas exchange velocity and carbon efflux using the SWOT Satellite"
author: "CB Brinkerhoff, CJ Gleason, PA Raymond, MH Harlan"
date: "`r Sys.Date()`"
bibliography: References.bib
output:
  word_document: 
      reference_docx: word-styles-reference.docx
---

@@

##1 Introduction
Greenhouse gas (GHG) emissions from rivers and lakes are significant [@colePlumbingGlobalCarbon2007a]. Rivers are usually supersaturated with dissolved GHGs and present a net flux of these gases from water to air [@coleCarbonCatchmentsConnecting2001a; @jonesLongtermDeclineCarbon2003]. This flux $[M/L^2T]$ is calculated, with in situ knowledge of the gas concentration gradient between the water $C_{water} [M/L^3]$ and the air $C_{air} [M/L^3]$ and the gas exchange velocity $k [L/T]$, as equation 1.

$$Flux=k(C_{water}-C_{air}) (1)$$

k is largely a function of water column turbulence, at least in fluvial settings. This is because k scales with diffusivity as normalized by the boundary-layer thickness, and in rivers turbulence defines the thickness of the boundary-layer (@hallGasExchangeStreams2020 for a thorough review of gas exchange velocity in rivers). Due to its reliance on diffusivity (as well as kinematic viscosity), k is defined differently for different GHGs [@raymondScalingGasTransfer2012a; @wanninkhofRelationshipWindSpeed1992a]. It is also often measured using different gases than those being studied, so a variant of k that is normalized to a Schmidt number of 600 ($k_{600} [L/T]$) is generally used to be GHG and temperature agnostic [@hallGasExchangeStreams2020]. This is synonymous with a k for carbon dioxide ($k_{CO2}$) at 20 degrees Celsius.

It is infeasible to measure $k_{600}$ directly when working at the river network scale and so ‘upscaling functions’ are used that predict $k_{600}$ from readily available river geomorphology [e.g. @oconnorMechanismReaerationNatural1958; @palumboAssessingPerformanceReaeration2014; @raymondScalingGasTransfer2012a]. Following the theory above, many upscaling functions are based on the basic concept that k is empirically correlated with the turbulent energy dissipation e at the air-water interface [@zappaEnvironmentalTurbulentMixing2007]. This is convenient for upscaling because in rivers and streams e can be equated to the turbulent energy dissipation rate eD $[L^2/T^3]$ [@ulsethDistinctAirWater2019a; @raymondScalingGasTransfer2012a], which is estimated using easily calculated hydraulic parameters (eD≈gSV where g is gravitational acceleration, S is channel slope, and V is average flow velocity- @tsivoglouTracerMeasurementReaeration1976). It has been observed that this scaling relationship is particularly strong in higher energy streams where bubble-mediated gas exchange occurs, though the general scaling function holds across all rivers and streams [@ulsethDistinctAirWater2019a]. Despite reasonable predictive accuracy, these upscaling models suffer from residual variation indicating large predictive error (Wallin et al. 2011). Still, these are the current best options for estimating $k_{600}$ in rivers. It has been suggested that exploring at-a-station hydraulic variation (and thus $k_{600}$ variation) is a promising approach to reducing the residual variation in upscaling models [@hallGasExchangeStreams2020].

Because channel slope is readily available in any hydrographic data product, most efforts to upscale $k_{600}$ to river networks are limited by estimates of the reach-averaged flow velocity of the river channel. Therefore, network-scale GHG flux estimates exist in some part at the mercy of the quality of average flow velocity data. Ideally, a worker can directly measure velocity at some permanently or temporarily gauged river in their system and build velocity~discharge rating curves to upscale to the entire river network. Rating curves are also referred to as hydraulic geometry (HG) models: simple power law relations between hydraulic properties and discharge [@leopoldHydraulicGeometryStream1953a]. However, this is not feasible in ungauged basins (whether physically or politically ungauged- @gleasonRemoteSensingRiver2020) where no in situ hydraulic information is available and fieldwork is impractical. Thus, in ungauged settings, present understandings of GHG fluxes from rivers are likely limited by our access to velocity and/or discharge data.

To circumvent this problem in ungauged basins/at massive spatial scales, researchers have used ‘global-scope’ HG models that use a single set of model parameters trained on large hydraulics datasets. This is the approach used in upscaling average flow velocity to continental/global river networks and has proven quite useful in those settings (e.g. @horgbyUnexpectedLargeEvasion2019; @borgesVariabilityGasTransfer2004; @lauerwaldSpatialPatternsCO22015; @raymondGlobalCarbonDioxide2013a). However, it has long been established that HG parameters exhibit high spatial variability across landscapes and vary significantly from river to river [@parkWorldwideVariationsHydraulic1977]. Because of this, a large body of geomorphology work has attempted to parse out process-based explanations for HG parameters (e.g. @dingmanAnalyticalDerivationAtastation2007b; @parkerPhysicalBasisQuasiuniversal2007b; @singhTHEORIESHYDRAULICGEOMETRY2003a; @fergusonHydraulicsHydraulicGeometry1986a) and an open research question is how to best prescribe a set of HG parameters to a given river (particularly at the global scale). A potential alternative to this approach, in ungauged basins at least, is to directly estimate a river’s hydraulic properties from remote sensing (RS) data.

The problem of hydraulic estimation in ungauged basins is itself a burgeoning subfield within fluvial hydrology and an analogue to the problem presented here is the ungauged remote sensing of river discharge (RSQ- see @gleasonRemoteSensingRiver2020 for a thorough review). A specific subset of these ungauged RSQ methods are termed ‘McFLIs’ or ‘Mass Conserved Flow Law Inversion’ algorithms (e.g. @andreadisConstrainingAssimilationSWOT2020d; @brinkerhoffConstrainingRemoteRiver2020; @durandEstimatingReachaveragedDischarge2014a; HIVDI; @hagemannBAMBayesianAMHGManning2017a). They use basic geomorphic theories, rather than full hydrologic and/or hydraulic models, and the concept of ‘prior river knowledge’ to estimate discharge from RS data where not all hydraulic terms are RS-able [@gleasonRemoteSensingRiver2020; @gleasonTrackingRiverFlows2017]. McFLIs have largely been developed in the context of the NASA/CNES/UKSA/CSA Surface Water and Ocean Topography (SWOT) mission, which will launch in 2022 and provide the world’s first global measurements of water surface extent and elevation at novel temporal resolutions [@biancamariaSWOTMissionIts2016]. On average SWOT will sample rivers every 11 days and provide a unique opportunity to measure river and lake hydraulics at the global scale. While the raw performance of McFLIs is lower than a more sophisticated approach, they are globally flexible, easily implementable in any river that SWOT can sample, and improve our understandings of ungauged rivers when little to no information was previously available [@brinkerhoffConstrainingRemoteRiver2020; @durandIntercomparisonRemoteSensing2016b]. Recently, McFLIs have also shown promise in providing additional and beneficial information to traditional hydrologic modeling via data assimilation [@ishitsukaCombiningOpticalRemote].

McFLIs often employ Bayesian inference. Bayesian techniques have been previously used to estimate $k_{600}$ concurrently with stream metabolism from dissolved oxygen (O2) datasets [@applingOvercomingEquifinalityLeveraging2018; @graceFastProcessingDiel2015; @holtgrieveSimultaneousQuantificationAquatic2010]. While this approach requires detailed field data and is thus not applicable to $k_{600}$ estimation in ungauged rivers, it does suggest that Bayesian inference could be useful for estimating $k_{600}$ when parameter equifinality is a problem (at it is with McFLIs).

Within this context, we sought to explore estimating $k_{600}$ in ungauged basins with no in situ information. We hypothesize that remote sensing $k_{600}$ is possible by combining 1) hydraulic geometry theory, 2) $k_{600}$ upscaling theory, 3) Bayesian inference (in a manner similar to McFLI), and 4) SWOT data. More specifically, this chapter aims to answer the following questions:
	1) Is upscaling the gas exchange velocity with no a priori knowledge of eD possible?
	2) Is ungauged RS of gas exchange velocity possible?
	3) How might this algorithm influence upscaling of riverine carbon efflux?

To answer these questions, we build a RS algorithm that ingests SWOT data and produces $k_{600}$ estimates (with fully quantified uncertainty within a Bayesian framework) that requires no in situ information about a river’s hydraulic properties. We validate the algorithm on simulated SWOT data for 22 rivers. Finally, we compare our algorithm against established methods: we use previously published carbon dioxide ($CO_2$) data to compare the bulk carbon efflux (via $CO_2$ evasion) from the 22 rivers as calculated using both our new algorithm and established velocity rating curves and in situ discharge records.

##2 Methods
We name the RS of $k_{600}$ algorithm BIKER, or the ‘Bayesian Inversion/Inference of the $k_{600}$ Rate’ algorithm. To build BIKER, two steps are taken: first, we develop a $k_{600}$ upscaling model that does not rely on a priori knowledge of eD (section 2.1) and then we use that upscaling model to build the actual RS algorithm (section 2.2). Following these methods, the validation setup is described (section 2.3), as is the comparison of estimated bulk carbon effluxes from a suite of average flow velocity models (section 2.4). A flowchart detailing the complete workflow is included (Figure 1).

@@

###2.1 Developing an RS-able $k_{600}$ upscaling model
[@raymondScalingGasTransfer2012a] first used river hydraulics and the theoretical relationship between the turbulent dissipation rate (eD) and gas exchange to upscale $k_{600}$ from average flow velocity and channel slope, extending earlier work predicting the reaeration coefficient (e.g. @oconnorMechanismReaerationNatural1958; @tsivoglouTracerMeasurementReaeration1976). Later, [@ulsethDistinctAirWater2019a] built upon [@raymondScalingGasTransfer2012a]’s work by expanding their dataset to include measurements from steeper rivers and finding that two distinct upscaling regimes exist in low and high energy rivers. More specifically, [@ulsethDistinctAirWater2019a] scale $k_{600}$ using eD, where the resulting upscaling model parameters are significantly different whether eD is high or low. Their upscaling function is reprinted as equation 2.

$$ln(k_{600})=3.10 + 0.35ln(eD) if eD < 0.02  (2) \\

ln(k_{600})=6.43+1.18ln(eD) if eD > 0.02$$

While equation 2 would be ideal for the purposes outlined for this manuscript and provides best-to-date predictive accuracy for upscaling $k_{600}$, it requires an a priori knowledge of eD to assign its differential model parameters. We do not have this a priori knowledge in an ungauged setting and only have access to SWOT-observed river widths and heights (and therefore water surface slopes). Thus, we need to extend the [@ulsethDistinctAirWater2019a] model to assign the two model parameters using only RS-able hydraulics. To do this, we use the basic premise that the $k_{600}\sim eD$ scaling relationship is different for different river sizes. We use river width as a proxy for river size and train different upscaling functions for different groups of data with similar river sizes (Figure 2, trained on the [@ulsethDistinctAirWater2019a] dataset). In doing so, we find that narrow rivers (< 10m wide) need to be further broken into low energy (slope < 0.05) and high energy (slope > 0.05) regimes in order to meet the assumptions necessary to fit a linear regression model (specifically, normally distributed residuals of the predictions).

@@

We combined these five regression models into a ‘rule-based regression model’ (equation 3) for river size i and model parameters a and b. Per Figure 2, there are five possible sets of a and b parameters, as defined for rivers 10-50m wide, 50-100m wide, 100+m wide, low energy rivers 0-10m wide (slope < 0.05), and high energy rivers 0-10m wide (slope > 0.05).

$$k_{600}=a_i(eD)^{b_i} (3)$$

This model therefore assigns ai and bi a priori using just river width and slope to define i. Another implication of this extension of the [@ulsethDistinctAirWater2019a] model is that we now use a and b parameters that implicitly reflect the distinct geomorphology of different sized rivers.

Finally, we tested whether this model can reasonably estimate $k_{600}$ relative to the original [@ulsethDistinctAirWater2019a] model, with the goal of at least replicating their performance. We compared it against the various eD-based upscaling models from [@raymondScalingGasTransfer2012a] that are commonly used in the literature. This was done by training all upscaling functions on 80% of the [@ulsethDistinctAirWater2019a] data and comparing their performance on the remaining 20% of the data. Because the [@raymondScalingGasTransfer2012a] models were originally trained on a subset of this dataset, we deemed it fairer to refit their parameters to this expanded dataset (but keep model structure identical).

###2.2 Developing the BIKER algorithm
With a validated upscaling model that requires no a priori knowledge of eD (results presented in section 3.1) we turn to developing the algorithm to RS $k_{600}$. The approach used here is heavily informed by [@hagemannBAMBayesianAMHGManning2017a]’s algorithm for ungauged RSQ via Bayesian inference. Consult that paper for the logic behind conceptualizing the equation for discharge as a Bayesian inference problem, which we largely follow here to conceptualize upscaling $k_{600}$ as a Bayesian inference problem.

@@

In order to use Bayesian inference to solve for $k_{600}$, we need to derive a sampling model/likelihood function. Following [@tsivoglouTracerMeasurementReaeration1976], $eD \approx gS_hV$ for unsteady flow conditions (channel slope $S_0 \neq S_h$- see text S1 for this model derivation and why assuming non-uniform flow is convenient for our application). Using this, as well as Manning’s equation to relate average flow velocity to channel geometry, we rewrite equation 3 as equation 4. In equation 4, we use the Manning’s formulation from Hagemann et al. (2017) which formulates depth via river width and area (in order to use SWOT-observables). $k_{600}$ is now defined as a function of $S_h$, channel roughness n $[T/L^{\frac{1}{3}}]$, channel width W [L], and channel area $A=A_0+dA [L^2]$ for the change in channel area dA). It is also defined by the upscaling parameters $a_i$ and $b_i$.

$$k_{600}=a_i(gS_h^{\frac{3}{2}}\frac{1}{n}(A_0+dA)^\frac{2}{3}W^\frac{-2}{3})^{b_i} (4)$$

@@

$$a_i^{-1}(gS_h^\frac{-3}{2}W^\frac{2}{3})^{b_i} \sim normal(k_{600}^{-1}\frac{1}{n}(A_0+dA)^\frac{2}{3})^{b_i}, \sigma^2_{upscale})  (5)$$

Equation five necessitates that Bayesian priors be assigned to model parameters n, $A_0$, and $k_{600}$. Bayesian priors formularize the a priori estimates (and uncertainties) for these parameters. Or more intuitively, they represent our ‘prior river knowledge’ of what n, A0, and $k_{600}$ probably are for some river given that they cannot be remotely sensed. All priors are formalized within the model as truncated normal distributions of the ln-transformed terms such that $ln(X) \sim N(\mu, \sigma^2)$ for λ<X<ε, using prior hyperparameters mean (μ), standard deviation (σ), and upper (ε) and lower bounds (λ) for parameter X = n, $A_0$, or $k_{600}$. In order to avoid relying on in situ information, we assign prior hyperparameters using SWOT data only. n and $A_0$ prior hyperparameters were assigned following the method developed by [@brinkerhoffConstrainingRemoteRiver2020]. They developed a set of river channel prior hyperparameters for McFLI algorithms that are entirely RS-able and reflect differential channel hydraulics as a function of river geomorphology. With this method, they showed substantial improvement in the accuracy of their ungauged RSQ estimates with no changes to the algorithms or data used. We assigned the $k_{600}$ prior hyperparameters using a simple slope regression model trained on the [@ulsethDistinctAirWater2019a] dataset to assign μ (equation 6). $σ^2$ was set to the standard error of equation 6: 1.023. λ and ε were set to -3 and 2.7, respectively.

$$\mu_{k600}=3.22+0.347ln(S_h) if ln(S_h) < -4.634   (6) \\

\mu_{k600}=6.85+1.13ln(S_h) if ln(S_h)>-4.634$$

@@

###2.3 BIKER validation setup
We validated the BIKER algorithm on 22 rivers using observed average flow velocity and observed slope. First, we detail the 22 rivers, and then we detail how ‘observed’ $k_{600}$ is calculated.

Because SWOT has yet to launch, it is standard practice to benchmark SWOT-related algorithms on ‘SWOT-like data’. There are three current types of SWOT-like data: 1) AirSWOT, which is an airborne Ka-band inSAR currently limited to five rivers globally, 2) simulated rivers that mimic the type of data SWOT will provide, and 3) the SWOT simulator, which introduces errors to these simulated rivers (mostly from radar layover error and random noise- @oubanasDischargeEstimationUngauged2018). Because we are principally interested in algorithm performance, we limit our validation setup to simulated rivers in order to benchmark across as many rivers as possible. These simulated rivers are simply reach-averaged hydraulic model outputs where the water surface heights and widths are labelled as ‘RS observations’ and are used as the sole inputs to the BIKER algorithm. Simulated rivers mimic perfect measurement conditions (i.e. no observation errors) and represent the best-case scenarios of what SWOT will provide to hydrologists. Here, we use the 22 rivers archived in [@frassonCompilationHydraulicModels2019a] and used by [@rodriguezObservingRiversVarying2020] to develop reach-averaged Saint Vernant equations within the SWOT context. We omitted the Arial Khan river in Bangladesh due to known problems with that hydraulic model and we further sampled the models for only the observations every 11 days to mimic the average SWOT overpass frequency (this also approximately algins the SWOT observations with the bi-weekly CO2 data used later- Section 2.4). All told, this yielded 503 sets of SWOT observations to validate BIKER with.

>>>>

---
title: "Remote sensing riverine gas exchange from SWOT and implications for estimating carbon efflux from rivers"
author: "CB Brinkerhoff, CJ Gleason, PA Raymond, MH Harlan"
date: "`r Sys.Date()`"
bibliography: References.bib
csl: american-geophysical-union.csl
output:
  word_document: 
      reference_docx: word-styles-reference.docx
---

@@

##Key Points
- We introduce the BIKER algorithm to remotely sense riverine gas exchange using the SWOT satellite and no in situ information
- BIKER can nearly replicate estimates of $CO_2$ fluxes made via hydraulic geometry and an in situ streamgauge
- Expected SWOT measurement errors exert a trivial influence on BIKER's estimates
- BIKER's accuracy is limited by process-level uncertainties in gas exchange from rivers and not by its algorithmic setup

##Abstract
Go BIKER!!
x
x
x
x
x
x
x
x
x
x

##1 Introduction
The inland water drainage network both transports and transforms various constiutents as they move from the landscape to the ocean and a byproduct of many of these processes is riverine greenhouse gas (GHG) production [@colePlumbingGlobalCarbon2007a]. Rivers are usually supersaturated with dissolved GHGs and present a significant flux of these gases from water to air [@raymondGlobalCarbonDioxide2013a]. This flux $[M/L^2T]$ is calculated, with in situ knowledge of the gas concentration gradient between the water $C_{water} [M/L^3]$ and the air $C_{air} [M/L^3]$ and the gas exchange velocity $k [L/T]$, as equation 1. Note that *k* is generally studied as normalized by a Schmidt number of 600 ($k_{600}$) to be gas and temperature agnostic [see @hallGasExchangeStreams2020 for a thorough review of gas exchange in rivers].

$$Flux=k(C_{water}-C_{air}) \text{(1)}$$

It is infeasible to measure $k_{600}$ directly when working at the drainage-network-scale and so ‘upscaling functions’ are used that predict $k_{600}$ from readily available river geomorphology [e.g. @oconnorMechanismReaerationNatural1958; @palumboAssessingPerformanceReaeration2014; @raymondScalingGasTransfer2012a]. Many upscaling functions are based on the basic theory that *k* is empirically correlated with the turbulent energy dissipation *e* at the air-water interface [@zappaEnvironmentalTurbulentMixing2007]. This is convenient for upscaling because in rivers and streams *e* can be equated to the turbulent energy dissipation rate *eD* $[L^2/T^3]$ [@ulsethDistinctAirWater2019a; @raymondScalingGasTransfer2012a], which is estimated using easily calculated hydraulic parameters: $eD \approx gSV$ where *g* is gravitational acceleration $[L/T^2]$, *S* is channel slope, and *V* is average flow velocity $[L/T]$ [@tsivoglouTracerMeasurementReaeration1976]. Because *S* is readily available in any hydrographic data product, most efforts to upscale $k_{600}$ to drainage networks are limited by estimates of the *V* of the river channel. Thus, drainage-network-scale GHG emissions estimates are sensitive to their *V* data. This is exasperated in ungauged basins where *V* models cannot be trained because no in situ hydraulic information is available and fieldwork is impractical [@gleasonRemoteSensingRiver2020].

To circumvent this problem in ungauged basins/at drainage-network-scales, to date researchers have used ‘global-scope’ hydraulic geometry (HG) models: simple power law relations between hydraulic properties and discharge [@leopoldHydraulicGeometryStream1953a] that are trained on extensive hydraulics datasets. This is the approach used in upscaling *V* to continental/global river networks and has proven useful in those settings [e.g. @horgbyUnexpectedLargeEvasion2019; @borgesVariabilityGasTransfer2004; @lauerwaldSpatialPatternsCO22015; @raymondGlobalCarbonDioxide2013a]. However, it has long been established that HG parameters exhibit high spatial variability across landscapes and vary significantly from river to river [@parkWorldwideVariationsHydraulic1977]. Because of this, a large body of geomorphology work has attempted to parse out process-based explanations for HG parameters [e.g. @dingmanAnalyticalDerivationAtastation2007b; @parkerPhysicalBasisQuasiuniversal2007b; @singhTHEORIESHYDRAULICGEOMETRY2003a; @fergusonHydraulicsHydraulicGeometry1986a] and an open research question is how to best prescribe a set of HG parameters to a given river (particularly at the global scale). A potential alternative to this approach is to directly estimate a river’s hydraulic properties using remote sensing (RS) data.

Remote hydraulic estimation is itself a burgeoning subfield within remote sensing of hydrology and an analogue to the problem presented here is the remote sensing of river discharge (RSQ- see @gleasonRemoteSensingRiver2020 for a thorough review). There are many ways to perform RSQ, from calibrating hydrologic and hydraulic models [] to methods that use no in situ information ('ungauged approaches') []. Many of these ungauged approaches were developed in the context of the upcoming NASA/CNES/UKSA/CSA Surface Water and Ocean Topography (SWOT) satellite mission.

SWOT will launch in 2022 and provide the world’s first global measurements of water surface extent and elevation at novel temporal resolutions. SWOT is a wide swath Ka-band radar interferometer and will sample rivers every 1 to 7 days per 21 day repeat cycle. It will measure rivers wider than 100m with a goal of expanding this to rivers at least 50m wide [@biancamariaSWOTMissionIts2016]. In the context of RSQ, a decade of SWOT work has explored the multi-parameter problem of estimating roughness and bathymetry to produce the SWOT discharge product [e.g. @andreadisConstrainingAssimilationSWOT2020d; @brinkerhoffConstrainingRemoteRiver2020; @durandEstimatingReachaveragedDischarge2014a; @garamboisInferenceEffectiveRiver2015; @gleasonRetrievalRiverDischarge2014a; @hagemannBAMBayesianAMHGManning2017a; @larnierRiverDischargeBathymetry2020a; @oubanasDischargeEstimationUngauged2018]. A specific subset of these ungauged RSQ methods are termed ‘McFLIs’ or ‘Mass Conserved Flow Law Inversion’ algorithms (e.g. @andreadisConstrainingAssimilationSWOT2020d; @brinkerhoffConstrainingRemoteRiver2020; @durandEstimatingReachaveragedDischarge2014a). They use basic geomorphic theories, rather than full hydrologic and/or hydraulic models, and the concept of ‘prior river knowledge’ to estimate discharge from RS data where not all hydraulic terms are RS-able [@gleasonRemoteSensingRiver2020; @gleasonTrackingRiverFlows2017]. While the raw performance of McFLIs is lower than the more sophisticated approaches mentioned above, they are globally flexible, easily implementable in any river that SWOT can sample, and improve our hydrological understandings of ungauged rivers when little to no information was previously available [@brinkerhoffConstrainingRemoteRiver2020; @durandIntercomparisonRemoteSensing2016b]. Recently, McFLIs have also shown promise in providing additional and beneficial information to traditional hydrologic modeling via data assimilation [@ishitsukaCombiningOpticalRemote2020].

McFLIs often employ Bayesian inference. Bayesian techniques have been previously used to estimate $k_{600}$ concurrently with stream metabolism from dissolved oxygen ($O_2$) datasets [@applingOvercomingEquifinalityLeveraging2018; @graceFastProcessingDiel2015; @holtgrieveSimultaneousQuantificationAquatic2010]. While this approach requires detailed field data and is thus not applicable to $k_{600}$ estimation in ungauged rivers, it does suggest that Bayesian techniques could be useful for estimating $k_{600}$ when parameter equifinality is a problem (as it is with McFLIs).

Thus, to advance our understanding of riverine GHG fluxes in ungauged basins and/or at drainage-network-scales, we sought to estimate $k_{600}$ using only RS data and no in situ information. We hypothesize that remote sensing $k_{600}$ is possible by coupling 1) hydraulic geometry theory, 2) $k_{600}$ upscaling theory, 3) Bayesian remote sensing (in a manner similar to McFLI), and 4) SWOT data. More specifically, this manuscript aims to answer two questions:
	1) Is ungauged RS of the gas exchange velocity possible?
	2) How might this algorithm influence upscaled estimates of riverine carbon efflux to the atmosphere?

To answer these questions, we build a RS algorithm that ingests SWOT data and produces $k_{600}$ estimates (with fully quantified uncertainty within a Bayesian framework) that requires no in situ information about a river’s hydraulic properties. We validate the algorithm on simulated SWOT data for 22 rivers (as SWOT has yet to launch, simulated data is necessary- Section 2.3). We also quantify the influence of the expected measurement errors in SWOT data on our $k_{600}$ estimates. Finally, we compare our algorithm against published methods: we use previously published dissolved carbon dioxide ($CO_2$) data to compare the bulk carbon efflux (via $CO_2$ evasion) from the 22 rivers as calculated using both our new algorithm and published velocity rating curves and in situ discharge records.

##2 Methods
We name the RS of $k_{600}$ algorithm BIKER, or the ‘Bayesian Inversion of the $k_{600}$ Rate’ algorithm. To build BIKER, two steps are taken: first, we develop a $k_{600}$ upscaling model that does not rely on a priori knowledge of the river (section 2.1) and then we use that upscaling model to build the actual RS algorithm (section 2.2). Following these methods, the validation setup is described (section 2.3), as is the comparison of estimated bulk carbon effluxes from a suite of average flow velocity models (section 2.4). A flowchart detailing the complete workflow is included (Figure 1).

@@

###2.1 Developing an RS-able $k_{600}$ upscaling model
[@raymondScalingGasTransfer2012a] first used river hydraulics and the theoretical relationship between *eD* and gas exchange to upscale $k_{600}$ from average flow velocity and channel slope, extending earlier work predicting the reaeration coefficient (e.g. @oconnorMechanismReaerationNatural1958; @tsivoglouTracerMeasurementReaeration1976). Later, [@ulsethDistinctAirWater2019a] built upon [@raymondScalingGasTransfer2012a]’s work by expanding their dataset to include measurements from steeper rivers and finding that two distinct upscaling regimes exist in low and high energy rivers. More specifically, [@ulsethDistinctAirWater2019a] scale $k_{600}$ using eD, where the resulting upscaling model parameters are significantly different whether *eD* is high or low. Their upscaling function is reprinted as equation 2.

$$ln(k_{600})=3.10 + 0.35ln(eD) if eD < 0.02  \text{(2)}$$
$$ln(k_{600})=6.43+1.18ln(eD) if eD > 0.02$$

While equation 2 would be ideal for the purposes outlined for this manuscript and provides best-to-date predictive accuracy for upscaling $k_{600}$, it requires an a priori knowledge of *eD* to assign its differential model parameters. We do not have this a priori knowledge in an ungauged setting and only have access to SWOT-observed river widths and heights (and therefore water surface slopes). Thus, we need to extend the [@ulsethDistinctAirWater2019a] model to assign the two model parameters using only RS-able hydraulics. To do this, we use the basic premise that the $k_{600}\sim eD$ scaling relationship is different for different river sizes. We use river width as a proxy for river size and train different upscaling functions for different groups of data with similar river sizes (Figure 2, trained on the [@ulsethDistinctAirWater2019a] dataset). In doing so, we find that narrow rivers (< 10m wide) need to be further broken into low energy (slope < 0.05) and high energy (slope > 0.05) regimes in order to meet the assumptions necessary to fit a linear regression model (specifically, normally distributed residuals of the predictions).

@@

We combined these five regression models into a ‘rule-based regression model’ (equation 3) for river size *i* and model parameters *a* and *b*. Per Figure 2, there are five possible sets of *a* and *b* parameters, as defined for rivers 10-50m wide, 50-100m wide, 100+m wide, low energy rivers 0-10m wide (slope < 0.05), and high energy rivers 0-10m wide (slope > 0.05).

$$k_{600}=a_i(eD)^{b_i} \text{(3)}$$

This model therefore assigns $a_i$ and $b_i$ a priori using just river width and slope to define *i*. Another implication of this extension of the [@ulsethDistinctAirWater2019a] model is that we now use *a* and *b* parameters that implicitly reflect the distinct geomorphology of different rivers.

Finally, we tested whether this model can reasonably estimate $k_{600}$ relative to the original [@ulsethDistinctAirWater2019a] model, with the goal of at least replicating their performance. We compared it against the various *eD*-based upscaling models from [@raymondScalingGasTransfer2012a] that are commonly used in the literature. This was done by training all upscaling functions on 80% of the [@ulsethDistinctAirWater2019a] data and comparing their performance on the remaining 20% of the data. Because the [@raymondScalingGasTransfer2012a] models were originally trained on a subset of this dataset, we deemed it fairer to refit their parameters to this expanded dataset (but keep model structure identical).

###2.2 Developing the BIKER algorithm
With a validated upscaling model that requires no a priori knowledge of *eD* (results presented in section 3.1) we turn to developing the algorithm to RS $k_{600}$. The approach used here is heavily informed by [@hagemannBAMBayesianAMHGManning2017a]’s algorithm for ungauged RSQ via Bayesian inference. Consult that paper for the logic behind conceptualizing the equation for discharge as a Bayesian inference problem, which we largely follow here to conceptualize upscaling $k_{600}$ as a Bayesian inference problem.

@@

In order to use Bayesian inference to solve for $k_{600}$, we need to derive a sampling model/likelihood function. Following [@tsivoglouTracerMeasurementReaeration1976], $eD \approx gS_hV$ for unsteady flow conditions (channel slope $S_0 \neq S_h$- see text S1 for this model derivation and why assuming non-uniform flow is convenient for our application). Using this, as well as Manning’s equation to relate average flow velocity to channel geometry, we rewrite equation 3 as equation 4. In equation 4, we use the Manning’s formulation from Hagemann et al. (2017) which formulates depth via river width and area (in order to use SWOT-observables). $k_{600}$ is now defined as a function of $S_h$, channel roughness n $[T/L^{\frac{1}{3}}]$, channel width *W* [L], and channel area $A=A_0+dA [L^2]$ for the change in channel area *dA*). It is also defined by the upscaling parameters $a_i$ and $b_i$.

$$k_{600}=a_i(gS_h^{\frac{3}{2}}\frac{1}{n}(A_0+dA)^\frac{2}{3}W^\frac{-2}{3})^{b_i} \text{(4)}$$

@@

$$a_i^{-1}(gS_h^\frac{-3}{2}W^\frac{2}{3})^{b_i} \sim normal(k_{600}^{-1}\frac{1}{n}(A_0+dA)^\frac{2}{3})^{b_i}, \sigma^2_{upscale})  \text{(5)}$$

Equation five necessitates that Bayesian priors be assigned to model parameters n, $A_0$, and $k_{600}$. Bayesian priors formularize the a priori estimates (and uncertainties) for these parameters. Or more intuitively, they represent our ‘prior river knowledge’ of what *n*, $A_0$, and $k_{600}$ probably are for some river given that they cannot be remotely sensed. All priors are formalized within the model as truncated normal distributions of the ln-transformed terms such that $ln(X) \sim N(\mu, \sigma^2)$ for λ<X<ε, using prior hyperparameters mean ($\mu$), standard deviation ($\sigma$), and upper (ε) and lower bounds (λ) for parameter X = n, $A_0$, or $k_{600}$. In order to avoid relying on in situ information, we assign prior hyperparameters using SWOT data only. *n* and $A_0$ prior hyperparameters were assigned following the method developed by [@brinkerhoffConstrainingRemoteRiver2020]. They developed a set of river channel prior hyperparameters for McFLI algorithms that are entirely RS-able and reflect differential channel hydraulics as a function of river geomorphology. With this method, they showed substantial improvement in the accuracy of their ungauged RSQ estimates with no changes to the algorithms or data used. We assigned the $k_{600}$ prior hyperparameters using a simple slope regression model trained on the [@ulsethDistinctAirWater2019a] dataset to assign $\mu$ (equation 6). $σ^2$ was set to the standard error of equation 6: 1.023. λ and ε were set to -3 and 2.7, respectively.

$$\mu_{k600}=3.22+0.347ln(S_h) if ln(S_h) < -4.634   \text{(6)}$$
$$\mu_{k600}=6.85+1.13ln(S_h) if ln(S_h)>-4.634$$

@@

###2.3 BIKER validation setup
We validated the BIKER algorithm on 22 rivers using observed $V$ and observed $S_h$ First, we detail the 22 rivers, and then we detail how ‘observed’ $k_{600}$ is calculated.

Because SWOT has yet to launch, it is standard practice to benchmark SWOT-related algorithms on ‘SWOT-like data’. There are three current types of SWOT-like data: 1) AirSWOT, which is an airborne Ka-band inSAR currently limited to five rivers globally, 2) simulated rivers that mimic the type of data SWOT will provide, and 3) the SWOT simulator, which introduces measurement errors to these simulated rivers. Because we are principally interested in algorithm performance, we limit our validation setup to simulated rivers in order to benchmark across as many rivers as possible. These simulated rivers are simply reach-averaged hydraulic model outputs where the water surface heights and widths are labelled as ‘RS observations’ and are used as the sole inputs to the BIKER algorithm. Simulated rivers mimic perfect measurement conditions (i.e. no measurement errors) and represent the best-case scenarios of what SWOT will provide to hydrologists. Here, we use the 22 rivers archived in [@frassonCompilationHydraulicModels2019a] and used by [@rodriguezObservingRiversVarying2020] to develop reach-averaged Saint Vernant equations within the SWOT context. We omitted the Arial Khan river in Bangladesh due to known problems with that hydraulic model and we further sampled the models for only the observations every 11 days to mimic the average SWOT overpass frequency (this also approximately algins the SWOT observations with the bi-weekly $CO_2$ data used later in this study- Section 2.4). All told, this yielded 503 sets of SWOT observations to validate BIKER with.

We also introduced measurement error into these simulated SWOT observations to assess the performance degradation due to errors inherent in SWOT measurements. Error in SWOT-measured river heights and water surface slopes will come from tolerances internal to the satellite, i.e. the SWOT error budget, as well as radar layover error. River width errors are assumed here to be neglible [@durandHowWillRadar2020a]. Layover error is the phenomenon when radar returns arrive at the sensor at the same time, generally in areas of high topographic variation [@reesPhysicalPrinciplesRemote2013]. While the SWOT simulator mentioned above is ideal for this analysis, it is cumbersome to run on many rivers and so here we coarsely model measurement error using the results from a recent global anlaysis of SWOT radar layover errors [@durandHowWillRadar2020a]. They developed a first-principles model for layover error from SWOT observations and, using global hydrography, quantified SWOT measurement error from both error tolerances and radar layover for 220,924 river reaches. The 68th percentiles of these reach-scale errors were 1.7cm/km and 10.4cm for water surface slope and height, respectively. To apply this model to our 22 rivers, we generate new river heights *H* and water surface slopes $S_h$ by sampling every observation from a normal distribution as defined as $S_{err_{i,t}} \sim normal(S_{obs_{i,t}}, 0.000017^{2})$ and $H_{err_{i,t}} \sim normal(H_{obs_{i,t}}, 0.104^{2})$ (in meters). Because BIKER ingests *dA*, rather than *H*, reach-scale height errors were converted to *dA* errors, again following [@durandHowWillRadar2020a]: $\sigma_{dA} \approx W\sigma_H\sqrt{2}$ given our rectangular channel assumptions.
